```{r}
library(Seurat)
#library(SeuratDisk)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(readxl)
library(scSorter)
library(patchwork)
library(HGC)
library(UCell)
```
```{r}
# set.seed(1)
# 
# sessionInfo()
# #remove.packages("SeuratObject")
# install.packages("SeuratObject")
```
```{r}
# Parameter
source("script/parameter.R")
```

```{r}
# import smartseq
smrt_mtx <- read.delim("data/neftel/IDHwtGBM.processed.SS2.logTPM.txt.gz")
rownames(smrt_mtx) <- smrt_mtx[, 1]
smrt_mtx <- smrt_mtx[, -1]
head(smrt_mtx)
colnames(smrt_mtx) <- gsub("\\.", "-", colnames(smrt_mtx))
smrt_mtx <- smrt_mtx[, order(colnames(smrt_mtx))]
head(smrt_mtx)

meta <- read.delim("data/neftel/metadata.txt")
meta <- meta[-1, ]
rownames(meta) <- meta$NAME
meta <- meta[order(rownames(meta)), ]
meta
library(Seurat)
smrt <- CreateSeuratObject(smrt_mtx, meta.data = meta,)

rm(smrt_mtx)
```
```{r}
# compare with original
library(readxl)
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 4)
signatures
```

```{r}
#Load 10x 
#x10_mtx <- read.delim("data/neftel/GSM3828673_10X_GBM_IDHwt_processed_TPM.tsv.gz")


#cannot use as the the 10x does not have cell annotation

```






#check data 
```{r}
# Check TPM
# As whether it is TPM yet`
hist(smrt$nCount_RNA)
```
```{r}
# Check non-malignant cell
unique(smrt$CellAssignment)
table(smrt$CellAssignment)
```

#Remove count and convert log2 to simple normalized
```{r}
# as the count is actually log TPM + 1 , we will move it to data layer and reverse the log to tpm
smrt[["RNA"]]$data <- smrt[["RNA"]]$counts
#smrt[["RNA"]]$counts <- NULL


#reverse the log2+1  and put to assay "norm"

# #code still bug
smrt[["norm"]] <- smrt[["RNA"]]
smrt[["norm"]]$data <- as((2^(smrt@assays$RNA$data)) - 1, "CsparseMatrix")
```


#QC, Normalization and scale
```{r}
library(tibble)
# filter by MT, Count, feature
Idents(smrt) <- "orig.ident"

# # as the smart seq does not need to do percent.mt
smrt[["percent.mt"]] <- PercentageFeatureSet(smrt, pattern = "^MT")


#rownames(smrt)[(grepl(pattern = "^MT",rownames(smrt),ignore.case = TRUE))]
```


```{r}
VlnPlot(smrt, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
# Percent.mt not found
```
#Subset only malignant cell assignment
```{r}
Idents(smrt) <- "CellAssignment"
smrt.mal <- subset(smrt, idents = "Malignant")
# rm(smrt)
```





```{r}
dim(smrt.mal)
```



```{r}
# Remove cell that have no score
colnames(smrt.mal@meta.data)


smrt.mal$isNA <- apply(smrt.mal@meta.data[c("MESlike2","MESlike1","NPClike1","NPClike2","AClike","OPClike")], 1, function(x) any(is.na(x)))
Idents(smrt.mal) <- "isNA"
smrt.mal <- subset(smrt.mal, idents = FALSE)

dim(smrt.mal)
```



#Label 
```{r}
smrt.mal.meta <- smrt.mal@meta.data

phenotype <- c("MESlike2", "MESlike1", "AClike", "OPClike", "NPClike1", "NPClike2")
score <- smrt.mal@meta.data[, phenotype]

score <- transform(score,
  MESlike2 = as.numeric(MESlike2),
  MESlike1 = as.numeric(MESlike1),
  AClike = as.numeric(AClike),
  OPClike = as.numeric(OPClike),
  NPClike1 = as.numeric(NPClike1),
  NPClike2 = as.numeric(NPClike2)
)


q10 <- apply(score, 2, function(x) {
  quantile(x, probs = c(0.10))
})

celltype.all <- c()
for (i in 1:nrow(score)) {
  score_cell <- sort(unlist((score[i, ])), decreasing = TRUE)
  score_cell
  MES <- c(score_cell["MESlike1"], score_cell["MESlike2"])
  min <- names(MES[which.min(MES)])

  score_cell <- score_cell[!names(score_cell) %in% min]

  NPC <- c(score_cell["NPClike1"], score_cell["NPClike2"])
  min <- names(NPC[which.min(NPC)])

  score_cell <- score_cell[!names(score_cell) %in% min]


  first <- score_cell[1]
  second <- score_cell[2]
  third <- score_cell[3]


  if ((second > 1) &
    (second > q10[names(second)]) &
    ((second - third) >= 0.3)) {
    # celltype <- paste0("Hybrid_",names(first),"_",names(second))
    celltype <- "Hybrid"
  } else {
    celltype <- names(score_cell[1])
  }
  celltype.all <- append(celltype.all, celltype)
}
```

```{r}
smrt.mal$celltype <- celltype.all
```

```{r}
table(smrt.mal$celltype == "Hybrid")
```



```{r}
# Merge MES and NPC
smrt.mal.meta <- smrt.mal@meta.data
smrt.mal.meta$celltype[(smrt.mal.meta$celltype == "MESlike1")] <- "MESlike"
smrt.mal.meta$celltype[(smrt.mal.meta$celltype == "MESlike2")] <- "MESlike"

smrt.mal.meta$celltype[(smrt.mal.meta$celltype == "NPClike1")] <- "NPClike"
smrt.mal.meta$celltype[(smrt.mal.meta$celltype == "NPClike2")] <- "NPClike"

smrt.mal$celltype_merge <- smrt.mal.meta$celltype
```






```{r}
# Skip log normalize as it is logTPM
smrt.mal <- FindVariableFeatures(smrt.mal)
smrt.mal <- ScaleData(smrt.mal,features = rownames(smrt.mal))
```


```{r}
smrt.mal <- RunPCA(smrt.mal, features = rownames(smrt.mal))
```

```{r}
ElbowPlot(smrt.mal)
```


```{r}
set.seed(123)
smrt.mal <- RunUMAP(smrt.mal, dims = 1:15)
smrt.mal <- RunTSNE(smrt.mal, dims = 1:15)
```

```{r}
DimPlot(smrt.mal, reduction = "umap", group.by = "Sample", label = FALSE)

DimPlot(smrt.mal, reduction = "tsne", group.by = "Sample")
```


```{r}
DimPlot(smrt.mal, reduction = "umap", group.by = "celltype", label = FALSE)
DimPlot(smrt.mal, reduction = "umap", group.by = "celltype_merge", label = FALSE)
```
```{r}
gc()
```

```{r}
cellstate_colors2 <- c("red","blue","yellow","purple4","purple1","purple3","green4","green1","green3")
names(cellstate_colors2) <- c("AClike","OPClike","Hybrid","MESlike1","MESlike2","MESlike","NPClike1","NPClike2","NPClike")
```


```{r}
tsne.celltype <- DimPlot(smrt.mal, reduction = "tsne", group.by = "celltype", label = FALSE,cols = cellstate_colors2) + labs(title = NULL) + theme(axis.text = element_blank(),axis.ticks = element_blank(),legend.position = "bottom",legend.text = element_text(size = 10),text = element_text(family = "Helvetica")) + guides(color = guide_legend(nrow = 1,override.aes = list(size=5)))

tsne.celltype
ggsave(plot = tsne.celltype,path = "figure/",filename = "cellstate.png",width = 8,height = 5)
```


```{r}
Idents(smrt.mal) <- "celltype"
smrt.mal.hybrid_rm <- subset(smrt.mal, idents = "Hybrid", invert = TRUE)
tsne.celltype_merge.noHybrid.merge <- DimPlot(smrt.mal.hybrid_rm, reduction = "tsne", group.by = "celltype_merge",cols = cellstate_colors2, label = FALSE)  + labs(title = NULL)+
  theme(axis.text = element_blank(),axis.ticks = element_blank(),legend.position = "right",legend.text = element_text(size = 10),text = element_text(family = "Helvetica"))

tsne.celltype_merge.noHybrid.merge
ggsave(plot = tsne.celltype_merge.noHybrid.merge,path = "figure/",filename = "cellstate_filter_merge.png",width = 8,height = 5)
```
```{r}
DimPlot(smrt.mal, reduction = "umap", group.by = "celltype_merge", label = FALSE, ncol = 7,split.by = "Sample",)
```


```{r}
# remove Hybrid
Idents(smrt.mal) <- "celltype"
smrt.mal <- subset(smrt.mal, idents = "Hybrid", invert = TRUE)
```

```{r}
smrt.mal$OPClike <- as.numeric(smrt.mal$OPClike)
# FeaturePlot(smrt.mal,features = "OPClike")


smrt.mal$NPClike1 <- as.numeric(smrt.mal$NPClike1)
# FeaturePlot(smrt.mal,features = "NPClike1")

smrt.mal$NPClike2 <- as.numeric(smrt.mal$NPClike2)
smrt.mal$MESlike2 <- as.numeric(smrt.mal$MESlike2)
smrt.mal$MESlike1 <- as.numeric(smrt.mal$MESlike1)
smrt.mal$AClike <- as.numeric(smrt.mal$AClike)
```


```{r}
FeaturePlot(smrt.mal,features = c("NPClike2","NPClike1","MESlike1","MESlike2","AClike","OPClike"))
```
```{r}
smrt.mal <- RunPCA(smrt.mal, features = unname(unlist(signatures)))
```

```{r}
ElbowPlot(smrt.mal)
```


```{r}
set.seed(123)
smrt.mal <- RunUMAP(smrt.mal, dims = 1:15)
smrt.mal <- RunTSNE(smrt.mal, dims = 1:15)
```

```{r}
 DimPlot(smrt.mal, reduction = "umap", group.by = "Sample", label = FALSE)
DimPlot(smrt.mal, reduction = "tsne", group.by = "Sample")

```


```{r}
DimPlot(smrt.mal, reduction = "umap", group.by = "celltype", label = FALSE)

DimPlot(smrt.mal, reduction = "umap", group.by = "celltype_merge", label = FALSE)
```
```{r}
DimPlot(smrt.mal, reduction = "umap", group.by = "celltype_merge", label = FALSE, ncol = 4,split.by = "Sample",)
```

```{r}
smrt.mal$OPClike <- as.numeric(smrt.mal$OPClike)
# FeaturePlot(smrt.mal,features = "OPClike")


smrt.mal$NPClike1 <- as.numeric(smrt.mal$NPClike1)
# FeaturePlot(smrt.mal,features = "NPClike1")

smrt.mal$NPClike2 <- as.numeric(smrt.mal$NPClike2)
smrt.mal$MESlike2 <- as.numeric(smrt.mal$MESlike2)
smrt.mal$MESlike1 <- as.numeric(smrt.mal$MESlike1)
smrt.mal$AClike <- as.numeric(smrt.mal$AClike)
```


```{r}
FeaturePlot(smrt.mal,features = c("NPClike2","NPClike1","MESlike1","MESlike2","AClike","OPClike"))
```



```{r}
#try norm instead of log-norm
smrt.mal.tpm <- smrt.mal
smrt.mal.tpm[["RNA"]]$scale.data <- NULL
smrt.mal.tpm[["RNA"]]$data <- as((2^(smrt.mal.tpm@assays$RNA@layers$data)) - 1, "CsparseMatrix")

```

```{r}
smrt.mal.tpm <- FindVariableFeatures(smrt.mal.tpm)
smrt.mal.tpm <- ScaleData(smrt.mal.tpm,features = rownames(smrt.mal.tpm))
smrt.mal.tpm <- RunPCA(smrt.mal.tpm, features = rownames(smrt.mal.tpm))
```

```{r}
ElbowPlot(smrt.mal.tpm)
```


```{r}
smrt.mal.tpm <- RunUMAP(smrt.mal.tpm, dims = 1:15)
smrt.mal.tpm <- RunTSNE(smrt.mal.tpm, dims = 1:15)
```

```{r}
p0 <- DimPlot(smrt.mal.tpm, reduction = "umap", group.by = "Sample", label = FALSE)
p0
```

```{r}
p0 <- DimPlot(smrt.mal.tpm, reduction = "tsne", group.by = "Sample")
p0
```


```{r}
p0 <- DimPlot(smrt.mal.tpm, reduction = "umap", group.by = "celltype", label = FALSE, split.by = "celltype", ncol = 4)
p0
```

#Build new geneset
```{r}
library(ggvenn)
# 
# if (!require(devtools)) install.packages("devtools")
# devtools::install_github("yanlinlin82/ggvenn")

# Load Dataset
smrt.mal <- readRDS("output/smrt_mal")

#assay RNA is logTPM
#assay norm is TPM
```


```{r}
smrt.mal@assays$RNA$data[1:10,1:10]
smrt.mal@assays$norm$data[1:10,1:10]
smrt@assays$RNA$counts[1:10,1:10]
smrt@assays$RNA$data[1:10,1:10]
```


# pseudobulk

```{r}
smrt.mal$celltype_merge <- as.factor(smrt.mal$celltype_merge)
```



```{r}
smrt.mal[["RNA"]]$counts <- smrt.mal[["RNA"]]$data

#aggregrate on logTPM
smrt.mal.pseudo <- AggregateExpression(smrt.mal,return.seurat = TRUE,group.by = c("Sample","celltype_merge"),assays = "RNA")

Idents(smrt.mal.pseudo) <- "celltype_merge"

bulk.deg.list <- list()
cellstates <-unique(smrt.mal.pseudo$celltype_merge)
bulk.deg.list <- list()
for (i in cellstates) {
  bulk.deg <- FindMarkers(object = smrt.mal.pseudo, 
                         ident.1 = i,
                       only.pos = TRUE)
  bulk.deg$fdr <- p.adjust(bulk.deg$p_val,method = "fdr")
  bulk.deg.list[[i]] <- bulk.deg
} 

# npc.deg <- FindMarkers(object = smrt.mal.pseudo, 
#                          ident.1 = "NPClike",
#                        only.pos = TRUE)
# opc.deg <- FindMarkers(object = smrt.mal.pseudo, 
#                          ident.1 = "OPClike",
#                        only.pos = TRUE)
# ac.deg <- FindMarkers(object = smrt.mal.pseudo, 
#                          ident.1 = "AClike",
#                        only.pos = TRUE)
# mes.deg <- FindMarkers(object = smrt.mal.pseudo, 
#                          ident.1 = "MESlike",
#                        only.pos = TRUE)



```



```{r}
#filter(avg_log2FC >1 & fdr < 0.05)  and slice top 50
bulk.deg.list.filter <- lapply(bulk.deg.list,function(x){
  x <- x %>%
    filter(avg_log2FC >1 & fdr < 0.05) %>% 
    arrange(fdr) %>%
    slice(1:50)
}) 
bulk.deg.list.filter
```


```{r}
bulk.deg.list.filter.genes <- lapply(bulk.deg.list.filter,rownames)
```

```{r}
vein <- ggvenn(bulk.deg.list.filter.genes)
vein <- vein + labs(title = "New")
vein
```
```{r}
# compare with original
library(readxl)
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 4)
colnames(signatures) <- paste0(colnames(signatures),"like")
```

```{r}
# Check intersection with each other
original <- list(NPClike = signatures$NPClike, AClike = signatures$AClike, MESlike = signatures$MESlike, OPClike = signatures$OPClike)

vein <- ggvenn(original)
vein <- vein + labs(title = "original")
vein
```


```{r}
venn_list <- lapply(cellstates, function(x){
  venn <- ggvenn(list(orig = signatures[,x,drop = TRUE], new = bulk.deg.list.filter.genes[[x]]), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = x)
  return(venn)
})
venn_list


```


```{r}
signatures.new <- readRDS(file = "output/signature_subtype_new")


signatures.new
```
```{r}
npc <- ggvenn(list(new = signatures.new$NPC.new, new.bulk = bulk.deg.list.filter.genes$NPClike,original =original$NPClike)) + labs(title = "NPC")

opc <- ggvenn(list(new = signatures.new$OPC.new, new.bulk = bulk.deg.list.filter.genes$OPClike,original =original$OPClike))+ labs(title = "OPC")

ac <- ggvenn(list(new = signatures.new$AC.new, new.bulk = bulk.deg.list.filter.genes$AClike,original =original$AClike))+ labs(title = "AC")

mes <- ggvenn(list(new = signatures.new$MES.new, new.bulk = bulk.deg.list.filter.genes$MESlike,original =original$MESlike))+ labs(title = "MES")

library(patchwork)

(npc + opc) / (ac + mes)
```

```{r}
#saveRDS(bulk.deg.list.filter.genes, file = "output/signature_subtype_bulk")
```

```{r}
library(stringi)


signatures.bulk.df <- data.frame(stri_list2matrix(bulk.deg.list.filter.genes))
colnames(signatures.bulk.df) <- names(bulk.deg.list.filter.genes)

signatures.bulk.df
```
```{r}
#write.csv(signatures.bulk.df, file = paste0("output/signature_subtype_", sig,"_","bulk", ".csv"), col.names = TRUE, row.names = FALSE)
```

```{r}
# Also compare to database
markers.db <- read_excel("data/Cell_marker_Human.xlsx")


markers.db.ac <- subset(markers.db, grepl("Astrocyte", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]

markers.db.ac <- subset(markers.db, cell_name == "Astrocyte")[, c(7, 9, 10)]




markers.db.mes <- subset(markers.db, grepl("Mesenchy", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]

markers.db.mes <- subset(markers.db, cell_name == "Mesenchymal cell")[, c(7, 9, 10)]

markers.db.npc <- rbind(subset(markers.db, grepl("neura", cell_name, ignore.case = TRUE))[, c(7, 9, 10)],
                        subset(markers.db, grepl("neuro", cell_name, ignore.case = TRUE))[, c(7, 9, 10)])
markers.db.npc <- subset(markers.db.npc, !grepl("non", cell_name, ignore.case = TRUE))
markers.db.npc <- subset(markers.db, cell_name == "Neural progenitor cell")[, c(7, 9, 10)]


markers.db.opc <- subset(markers.db, grepl("oligodendro", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
markers.db.opc <- subset(markers.db, cell_name == "Oligodendrocyte progenitor cell")[, c(7, 9, 10)]


markers.db.gbm <- subset(markers.db, grepl("glioblasto", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
unique(markers.db.gbm$cell_name)

marker.all <- list()
# marker.all <- append(marker.all, list(ac = unique(markers.db.ac$cell_name)))
# marker.all <- append(marker.all, list(mes = unique(markers.db.mes$cell_name)))
# marker.all <- append(marker.all, list(npc = unique(markers.db.npc$cell_name)))
# marker.all <- append(marker.all, list(opc = unique(markers.db.opc$cell_name)))
# marker.all <- append(marker.all,list(gbm = unique(markers.db.gbm$cell_name)))

marker.all <- append(marker.all, list(AClike = unique(markers.db.ac$Symbol)))
marker.all <- append(marker.all, list(MESlike = unique(markers.db.mes$Symbol)))
marker.all <- append(marker.all, list(NPClike = unique(markers.db.npc$Symbol)))
marker.all <- append(marker.all, list(OPClike = unique(markers.db.opc$Symbol)))
marker.all <- append(marker.all,list(GBM = unique(markers.db.gbm$Symbol)))

marker.all
```








```{r}
mes <- ggvenn(list(orig = signatures$MES, new = rownames(mes.de.other.markers), db = unique(markers.db.mes$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes

opc <- ggvenn(list(orig = signatures$OPC, new = rownames(opc.de.other.markers), db = unique(markers.db.opc$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "OPClike")
opc

npc <- ggvenn(list(orig = signatures$NPC, new = rownames(npc.de.other.markers), db = unique(markers.db.npc$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc

ac <- ggvenn(list(orig = signatures$AC, new = rownames(ac.de.other.markers), db = unique(markers.db.ac$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

(mes + opc) / (npc + ac)
```

```{r}
mes <- ggvenn(list(orig = signatures$MES, new = rownames(mes.de.other.markers), db = unique(markers.db.mes$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes

opc <- ggvenn(list(orig = signatures$OPC, new = rownames(opc.de.other.markers), db = unique(markers.db.opc$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "OPClike")
opc

npc <- ggvenn(list(orig = signatures$NPC, new = rownames(npc.de.other.markers), db = unique(markers.db.npc$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc

ac <- ggvenn(list(orig = signatures$AC, new = rownames(ac.de.other.markers), db = unique(markers.db.ac$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

(mes + opc) / (npc + ac)
```

# consistent among all donor
```{r}
Idents(smrt.mal) <- "Sample"

smrt.mal.list <- SplitObject(smrt.mal, split.by = "Sample")

celltype_all <- c("celltype_merge")
```


```{r}
findMarkersWrapper <- function(seurat_object, identity,celltype_chosen) {
  Idents(seurat_object) <- celltype_chosen
  
  tryCatch({
    FindMarkers(seurat_object, ident.1 = identity, only.pos = TRUE, min.cells.group = 1, min.cells.feature = 1)
  }, error = function(e) {
    # Optional: Print the error message
    print(paste("Error in processing", identity, ":", e$message))
    NULL  # Return NULL in case of error
  })
}
npc.de.other.markers.list <- lapply(smrt.mal.list, function(x) findMarkersWrapper(x, "NPClike",celltype_all[1]))
opc.de.other.markers.list <- lapply(smrt.mal.list, function(x) findMarkersWrapper(x, "OPClike",celltype_all[1]))
ac.de.other.markers.list <- lapply(smrt.mal.list, function(x) findMarkersWrapper(x, "AClike",celltype_all[1]))
mes.de.other.markers.list <- lapply(smrt.mal.list, function(x) findMarkersWrapper(x, "MESlike",celltype_all[1]))
```


```{r}
# library(tibble)
# 
# 
# npc.de.other.markers.list <- Filter(Negate(is.null), npc.de.other.markers.list)
# npc.list <- lapply(npc.de.other.markers.list,function(x) {
#   x <- rownames_to_column(x,var = "gene_name")
#   x$fdr <- p.adjust(x$p_val,method = "BH")
#   
#   return(x)
# })
# 
# 
# npc.de.other.markers.list <- Filter(Negate(is.null), npc.de.other.markers.list)
# npc.list <- lapply(npc.de.other.markers.list,function(x) {
#   x <- rownames_to_column(x,var = "gene_name")
#   x$fdr <- p.adjust(x$p_val,method = "BH")
#   
#   return(x)
# })
# 
# npc_df <- Reduce(rbind,npc.list)
# 
# threshold <- 0.3
# 
# npc_df_filtered <- npc_df %>%
#   filter(avg_log2FC >= 1,fdr < 0.05) %>% 
#   group_by(gene_name) %>%
#   mutate(count = n()) %>%
#   filter(count >= threshold*length(npc.list)) %>%
#   ungroup() %>%
#   select(-count)  # Optionally remove the count column

```

#original build genes
```{r}
# npc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "OPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# ac.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "MESlike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
npc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPClike", only.pos = TRUE)
opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "OPClike", only.pos = TRUE)
ac.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AClike", only.pos = TRUE)
mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "MESlike", only.pos = TRUE)

```


```{r}
npc.de.other.markers <- npc.de.other.markers[order(npc.de.other.markers$avg_log2FC, decreasing = TRUE), ]
npc.de.other.markers <- npc.de.other.markers[1:50, ]

opc.de.other.markers <- opc.de.other.markers[order(opc.de.other.markers$avg_log2FC, decreasing = TRUE), ]
opc.de.other.markers <- opc.de.other.markers[1:50, ]

ac.de.other.markers <- ac.de.other.markers[order(ac.de.other.markers$avg_log2FC, decreasing = TRUE), ]
ac.de.other.markers <- ac.de.other.markers[1:50, ]

mes.de.other.markers <- mes.de.other.markers[order(mes.de.other.markers$avg_log2FC, decreasing = TRUE), ]
mes.de.other.markers <- mes.de.other.markers[1:50, ]
# logfc.threshold and min.diff.pct are very arbitrary
```



```{r}
x <- list(NPC = rownames(npc.de.other.markers), AC = rownames(ac.de.other.markers), MES = rownames(mes.de.other.markers), OPC = rownames(opc.de.other.markers))


vein <- ggvenn(x)
vein <- vein + labs(title = "New")
vein
```


```{r}
# Check intersection with each other
x <- list(NPC = signatures$NPC, AC = signatures$AC, MES = signatures$MES, OPC = signatures$OPC)

vein <- ggvenn(x)
vein <- vein + labs(title = "original")
vein
```


```{r}
mes <- ggvenn(list(orig = signatures$MES, new = rownames(mes.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes

opc <- ggvenn(list(orig = signatures$OPC, new = rownames(opc.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "OPClike")
opc

npc <- ggvenn(list(orig = signatures$NPC, new = rownames(npc.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc

ac <- ggvenn(list(orig = signatures$AC, new = rownames(ac.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

library(patchwork)
(mes | opc) / (npc | ac)
```


```{r}
# Also compare to database
markers.db <- read_excel("data/Cell_marker_Human.xlsx")


markers.db.ac <- subset(markers.db, grepl("Astrocyte", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
unique(markers.db.ac$cell_name)
markers.db.ac <- subset(markers.db, cell_name == "Astrocyte")[, c(7, 9, 10)]




markers.db.mes <- subset(markers.db, grepl("Mesenchy", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
unique(markers.db.mes$cell_name)
markers.db.mes <- subset(markers.db, cell_name == "Mesenchymal cell")[, c(7, 9, 10)]

markers.db.npc <- rbind(subset(markers.db, grepl("neura", cell_name, ignore.case = TRUE))[, c(7, 9, 10)], subset(markers.db, grepl("neuro", cell_name, ignore.case = TRUE))[, c(7, 9, 10)])
markers.db.npc <- subset(markers.db.npc, !grepl("non", cell_name, ignore.case = TRUE))
unique(markers.db.npc$cell_name)
markers.db.npc <- subset(markers.db, cell_name == "Neural progenitor cell")[, c(7, 9, 10)]


markers.db.opc <- subset(markers.db, grepl("oligodendro", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
unique(markers.db.opc$cell_name)
markers.db.opc <- subset(markers.db, cell_name == "Oligodendrocyte progenitor cell")[, c(7, 9, 10)]


markers.db.gbm <- subset(markers.db, grepl("glioblasto", cell_name, ignore.case = TRUE))[, c(7, 9, 10)]
unique(markers.db.gbm$cell_name)

marker.all <- list()
marker.all <- append(marker.all, list(ac = unique(markers.db.ac$cell_name)))
marker.all <- append(marker.all, list(mes = unique(markers.db.mes$cell_name)))
marker.all <- append(marker.all, list(npc = unique(markers.db.npc$cell_name)))
marker.all <- append(marker.all, list(opc = unique(markers.db.opc$cell_name)))
# marker.all <- append(marker.all,list(gbm = unique(markers.db.gbm$cell_name)))
marker.all
```
```{r}
signatures$NPC
rownames(npc.de.other.markers)
unique(markers.db.npc$Symbol)

intersect(unique(markers.db.npc$Symbol), signatures$NPC)
Reduce(intersect, list(signatures$NPC, rownames(npc.de.other.markers), unique(markers.db.npc$Symbol)))
```




```{r}
mes <- ggvenn(list(orig = signatures$MES, new = rownames(mes.de.other.markers), db = unique(markers.db.mes$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes

opc <- ggvenn(list(orig = signatures$OPC, new = rownames(opc.de.other.markers), db = unique(markers.db.opc$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "OPClike")
opc

npc <- ggvenn(list(orig = signatures$NPC, new = rownames(npc.de.other.markers), db = unique(markers.db.npc$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc

ac <- ggvenn(list(orig = signatures$AC, new = rownames(ac.de.other.markers), db = unique(markers.db.ac$Symbol), gbm = unique(markers.db.gbm$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

(mes + opc) / (npc + ac)
```

```{r}
mes <- ggvenn(list(orig = signatures$MES, new = rownames(mes.de.other.markers), db = unique(markers.db.mes$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes

opc <- ggvenn(list(orig = signatures$OPC, new = rownames(opc.de.other.markers), db = unique(markers.db.opc$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "OPClike")
opc

npc <- ggvenn(list(orig = signatures$NPC, new = rownames(npc.de.other.markers), db = unique(markers.db.npc$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc

ac <- ggvenn(list(orig = signatures$AC, new = rownames(ac.de.other.markers), db = unique(markers.db.ac$Symbol)), set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

(mes + opc) / (npc + ac)
```

```{r}
# Add metamodule based on new score
signatures.new <- list()
signatures.new <- append(signatures.new, list(NPC.new = rownames(npc.de.other.markers)))
signatures.new <- append(signatures.new, list(OPC.new = rownames(opc.de.other.markers)))
signatures.new <- append(signatures.new, list(AC.new = rownames(ac.de.other.markers)))
signatures.new <- append(signatures.new, list(MES.new = rownames(mes.de.other.markers)))
signatures.new
```

```{r}
#saveRDS(signatures.new, file = "output/signature_subtype_new")
```

```{r}
library(stringi)
data.frame(stri_list2matrix(signatures.new))

signatures.new.df <- data.frame(stri_list2matrix(signatures.new))
colnames(signatures.new.df) <- names(signatures.new)

signatures.new.df
```
```{r}
#write.csv(signatures.new.df, file = paste0("output/signature_subtype_", sig, ".csv"), col.names = TRUE, row.names = FALSE)
```



#merge NPC/OPC then make new gene set

```{r}
library(ggvenn)

if (!require(devtools)) install.packages("devtools")
devtools::install_github("yanlinlin82/ggvenn")

# Load Dataset
smrt.mal <- readRDS("output/smrt_mal")
```



```{r}
smrt.mal@meta.data$celltype_npc_opc <- as.character(smrt.mal$celltype_merge)
smrt.mal@meta.data[smrt.mal$celltype_merge %in% c("NPClike","OPClike"),"celltype_npc_opc"] <- "NPC_OPClike"

smrt.mal$celltype_npc_opc
Idents(smrt.mal) <- "celltype_npc_opc"
```






```{r}
# npc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "OPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# ac.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "MESlike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)

npc_opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPC_OPClike", only.pos = TRUE)
ac.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AClike", only.pos = TRUE)
mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "MESlike", only.pos = TRUE)


npc_opc.de.other.markers <- npc_opc.de.other.markers[order(npc_opc.de.other.markers$avg_log2FC, decreasing = TRUE), ]
npc_opc.de.other.markers <- npc_opc.de.other.markers[1:50, ]



ac.de.other.markers <- ac.de.other.markers[order(ac.de.other.markers$avg_log2FC, decreasing = TRUE), ]
ac.de.other.markers <- ac.de.other.markers[1:50, ]

mes.de.other.markers <- mes.de.other.markers[order(mes.de.other.markers$avg_log2FC, decreasing = TRUE), ]
mes.de.other.markers <- mes.de.other.markers[1:50, ]
# logfc.threshold and min.diff.pct are very arbitrary
```



```{r}
x <- list(NPC_OPC = rownames(npc_opc.de.other.markers), AC = rownames(ac.de.other.markers), MES = rownames(mes.de.other.markers))


vein <- ggvenn(x)
vein <- vein + labs(title = "New")
vein
```
```{r}
# compare with original
library(readxl)
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 4)
signatures <- read.csv("output/signature_subtype_4metamodules_new_50.csv")
signatures
```

```{r}
# Check intersection with each other
x <- list(NPC = signatures$NPC.new, AC = signatures$AC.new, MES = signatures$MES.new, OPC = signatures$OPC.new)

vein <- ggvenn(x)
vein <- vein + labs(title = "original")
vein
```


```{r}
mes <- ggvenn(list(MES= signatures$MES.new, MES.npc_opc = rownames(mes.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "MESlike")
mes



npc_opc <- ggvenn(list(NPC = signatures$NPC.new,OPC= signatures$OPC.new ,NPC_OPC = rownames(npc_opc.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPClike")
npc_opc

ac <- ggvenn(list(AC = signatures$AC.new, AC.npc_opc = rownames(ac.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AClike")
ac

library(patchwork)
(mes | npc_opc | ac)
```

```{r}
# Add metamodule based on new score
signatures.new <- list()
signatures.new <- append(signatures.new, list(NPC_OPC = rownames(npc_opc.de.other.markers)))
signatures.new <- append(signatures.new, list(AC = rownames(ac.de.other.markers)))
signatures.new
```

```{r}
#saveRDS(signatures.new, file = "output/signature_subtype_npc_opc")
```

```{r}
library(stringi)
data.frame(stri_list2matrix(signatures.new))

signatures.new.df <- data.frame(stri_list2matrix(signatures.new))
colnames(signatures.new.df) <- names(signatures.new)

signatures.new.df
```
```{r}
sigs <- c( "4metamodules_new_50_mes","2metamodules_NPC_OPC", "3metamodules_new_50", "4metamodules_new_50")
sig <- sigs[2]

#write.csv(signatures.new.df, file = paste0("output/signature_subtype_", sig, ".csv"), col.names = TRUE, row.names = FALSE)

#read.csv(file = paste0("output/signature_subtype_", sig, ".csv"))
```









#merge NPC/OPC vs AC vs MES

```{r}
#No need for ref based
```



```{r}
#marker based

sigs <- c( "4metamodules_new_50_mes","2metamodules_NPC_OPC", "3metamodules_new_50", "4metamodules_new_50","npcopc_ac_mes")
sig <- sigs[4]
signatures.new.df_4 <- read.csv(file = paste0("output/signature_subtype_", sig, ".csv"))
signatures.new.df_4

sig <- sigs[2]
signatures.new.df_2 <- read.csv(file = paste0("output/signature_subtype_", sig, ".csv"))
signatures.new.df_2

signatures.new.df_5 <- cbind(signatures.new.df_2,signatures.new.df_4[,"MES.new",drop = FALSE])
colnames(signatures.new.df_5)[3] <- "MES"
signatures.new.df_5

sig <- sigs[5]
#write.csv(signatures.new.df_5, file = paste0("output/signature_subtype_", sig, ".csv"), row.names = FALSE)
```
#merge NPC/OPC vs AC/MES

```{r}
smrt.mal <- readRDS("output/smrt_mal")
```



```{r}
smrt.mal@meta.data$npcopc_acmes <- as.character(smrt.mal$celltype_merge)
smrt.mal@meta.data[smrt.mal$celltype_merge %in% c("NPClike","OPClike"),"npcopc_acmes"] <- "NPC_OPClike"
smrt.mal@meta.data[smrt.mal$celltype_merge %in% c("AClike","MESlike"),"npcopc_acmes"] <- "AC_MESlike"
Idents(smrt.mal) <- "npcopc_acmes"
```

```{r}
##saveRDS(smrt.mal, "output/smrt_mal", compress = TRUE)
```





```{r}
# npc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "OPClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# ac.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AClike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)
# mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "MESlike",logfc.threshold = 1,min.diff.pct = 0.1,only.pos = TRUE)

npc_opc.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "NPC_OPClike", only.pos = TRUE)
ac_mes.de.other.markers <- FindMarkers(smrt.mal, ident.1 = "AC_MESlike", only.pos = TRUE)


npc_opc.de.other.markers <- npc_opc.de.other.markers[order(npc_opc.de.other.markers$avg_log2FC, decreasing = TRUE), ]
npc_opc.de.other.markers <- npc_opc.de.other.markers[1:50, ]

ac_mes.de.other.markers <- ac_mes.de.other.markers[order(ac_mes.de.other.markers$avg_log2FC, decreasing = TRUE), ]
ac_mes.de.other.markers <- ac_mes.de.other.markers[1:50, ]

# logfc.threshold and min.diff.pct are very arbitrary
```



```{r}
x <- list(NPC_OPC = rownames(npc_opc.de.other.markers), AC_MES = rownames(ac_mes.de.other.markers))

library(ggvenn)
vein <- ggvenn(x)
vein <- vein + labs(title = "New")
vein
```
```{r}
# compare with original
library(readxl)
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 4)
signatures <- read.csv("output/signature_subtype_4metamodules_new_50.csv")
signatures
```

```{r}
# Check intersection with each other
x <- list(NPC = signatures$NPC.new, AC = signatures$AC.new, MES = signatures$MES.new, OPC = signatures$OPC.new)

vein <- ggvenn(x)
vein <- vein + labs(title = "original")
vein
```


```{r}



npc_opc <- ggvenn(list(NPC = signatures$NPC.new,OPC= signatures$OPC.new ,NPC_OPC = rownames(npc_opc.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "NPC_OPClike")


ac_mes <- ggvenn(list(AC = signatures$AC.new,MES= signatures$MES.new ,AC_MES = rownames(ac_mes.de.other.markers)), auto_scale = FALSE, set_name_size = 4, show_percentage = FALSE, text_size = 4) + labs(title = "AC_MESlike")

all_npc_opc <- venn::venn(list(NPC = signatures$NPC.new,OPC= signatures$OPC.new, AC = signatures$AC.new,MES= signatures$MES.new  , NPC_OPC = rownames(npc_opc.de.other.markers)))

all_ac_mes <- venn::venn(list(NPC = signatures$NPC.new,OPC= signatures$OPC.new, AC = signatures$AC.new,MES= signatures$MES.new, AC_MES = rownames(ac_mes.de.other.markers)))




library(patchwork)
(ac_mes | npc_opc)
all_npc_opc
all_ac_mes
```

```{r}
# Add metamodule based on new score
signatures.new <- list()
signatures.new <- append(signatures.new, list(NPC_OPC = rownames(npc_opc.de.other.markers)))
signatures.new <- append(signatures.new, list( AC_MES = rownames(ac_mes.de.other.markers)))
signatures.new
```

```{r}
##saveRDS(signatures.new, file = "output/signature_cellstate_npcopc_acmes")
```

```{r}
library(stringi)
data.frame(stri_list2matrix(signatures.new))

signatures.new.df <- data.frame(stri_list2matrix(signatures.new))
colnames(signatures.new.df) <- names(signatures.new)

signatures.new.df
```
```{r}
#marker based

sigs <- c( "4metamodules_new_50_mes","2metamodules_NPC_OPC", "3metamodules_new_50", "4metamodules_new_50","npcopc_ac_mes","npcopc_acmes")
sig <- sigs[6]

#write.csv(signatures.new.df, file = paste0("output/signature_subtype_", sig, ".csv"), col.names = TRUE, row.names = FALSE)

#read.csv(file = paste0("output/signature_subtype_", sig, ".csv"))
```









#archeive

```{r}
# Add metamodule based on new score
smrt.mal <- AddModuleScore_UCell(smrt.mal, features = signatures.new)
```

```{r}
smrt.mal@meta.data
```

```{r}
Idents(smrt.mal) <- "celltype_merge"
p0 <- FeaturePlot(smrt.mal, features = c("NPC.new_UCell", "OPC.new_UCell", "MES.new_UCell", "AC.new_UCell"))
p0
```


```{r}
VlnPlot(smrt.mal, features = c("NPC.new_UCell", "OPC.new_UCell", "MES.new_UCell", "AC.new_UCell"), ncol = 2)
```


```{r}
# Try nebula
# install.packages("devtools")
# library(devtools)
# install_github("lhe17/nebula")

library(nebula)
```



```{r}
##saveRDS(smrt.mal, "output/smrt_mal", compress = TRUE)
```


#feature plot sub-type signature and senescence
```{r}
# load subtpye signature
# signatures <- read_xlsx("data/gbm_subtype_genelist.xlsx",sheet = 2)
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 3)

colnames(signatures)[5] <- "G1_S"
colnames(signatures)[6] <- "G2_M"

signatures <- as.list(signatures)
signatures <- lapply(signatures, na.omit)

signatures <- signatures[-c(5, 6)]

signatures
```

```{r}
# smrt <- AddModuleScore(smrt, features =signatures,ctrl = 80, name = names(signatures))
smrt.mal <- AddModuleScore_UCell(smrt.mal, features = signatures)
```

```{r}
p0 <- DimPlot(smrt.mal, reduction = "umap", group.by = "celltype_merge", label = TRUE)
p0
```


```{r}
colnames(smrt.mal@meta.data)
```

```{r}
Idents(smrt.mal) <- "celltype_merge"
p0 <- FeaturePlot(smrt.mal, features = c("MES_UCell", "AC_UCell", "OPC_UCell", "NPC_UCell"), label = TRUE, repel = TRUE)
p0
```

```{r}
VlnPlot(smrt.mal, features = c("MES_UCell", "AC_UCell", "OPC_UCell", "NPC_UCell"), ncol = 2, )
```



#Senescence Lucy
```{r}
# create senescence score
signatures_sen <- list(senescence = scan("data/lucy_senesence_genes.txt", character(), sep = ",", strip.white = TRUE))
signatures_sen
```
```{r}
smrt.mal <- AddModuleScore(smrt.mal, features = signatures_sen, ctrl = 80, name = names(signatures_sen), search = TRUE)
```

```{r}
Idents(smrt.mal) <- "celltype_merge"
p0 <- FeaturePlot(smrt.mal, features = c("senescence1"), label = TRUE, repel = TRUE)
p0
```

```{r}
p0 <- VlnPlot(
  smrt.mal,
  features = "senescence1",
  group.by = "Sample",
)

p0 <- p0 + theme(axis.text.x.bottom = element_text(angle = 90, size = 5))
p0
```


```{r}
p0 <- VlnPlot(
  smrt.mal,
  features = "senescence1",
  group.by = "celltype_merge"
)

p0
```


```{r}
library(ggplot2)
# Basic box plot
p <- ggplot(smrt.mal@meta.data, aes(x = celltype_merge, y = senescence1, fill = celltype_merge)) +
  geom_boxplot()
p <- p + theme(
  axis.text.x.bottom = element_text(angle = 90, size = 5)
)
p
```


```{r}
library(ggplot2)
# Basic box plot
p <- ggplot(smrt.mal@meta.data, aes(x = celltype_merge, y = senescence1, fill = celltype_merge)) +
  geom_boxplot(outlier.shape = NA)
p <- p + facet_wrap(~Sample)
p <- p + theme(
  axis.text.x.bottom = element_text(angle = 90, size = 5)
)
p
```

```{r}
# Check whether it is statistically different
## Do ANOVA
aov.model <- aov(senescence1 ~ celltype_merge, data = smrt.mal@meta.data)
summary(aov.model)
```

```{r}
# create Q-Q plot to compare this dataset to a theoretical normal distribution
qqnorm(smrt.mal$senescence1, pch = 1, frame = FALSE)
```

```{r}
hist(smrt.mal$senescence1)
# Seem to have normal dis
```

```{r}
ks.test(smrt.mal$senescence1, "pnorm")
# No >>  distribution is not the same
```


```{r}
bartlett.test(senescence1 ~ celltype_merge, data = smrt.mal@meta.data)
# So there is heteroscedasticity
```


```{r}
# Kruskal-Wallis H test
krustal.model <- kruskal.test(senescence1 ~ celltype_merge, data = smrt.mal@meta.data)
krustal.model
```

```{r}
# pairwise.wilcox
pairwise.wilcox.test(smrt.mal$senescence1, smrt.mal$celltype_merge,
  p.adjust.method = "BH"
)
```
```{r}
unique(smrt.mal$celltype_merge)
```

```{r}
# Compare MES to the rest
MES <- data.frame(matrix(nrow = nrow(smrt.mal@meta.data), ncol = 1))
colnames(MES) <- "MESlike"
smrt.mal$MES <- MES
smrt.mal$MES[smrt.mal$celltype_merge == "MESlike"] <- "MESlike"
smrt.mal$MES[!smrt.mal$celltype_merge == "MESlike"] <- "others"

smrt.mal$MES <- as.factor(smrt.mal$MES)
```
```{r}
library(ggplot2)
# Basic box plot
p <- ggplot(smrt.mal@meta.data, aes(x = MES, y = senescence1, fill = MES)) +
  geom_boxplot()
p
```





```{r}
bartlett.test(senescence1 ~ MES, data = smrt.mal@meta.data)
# So there is heteroscedasticity
```
```{r}
ks.test(smrt.mal$senescence1, "pnorm")

# No >>  distribution is not the same
```



```{r}
model.wilcox <- wilcox.test(senescence1 ~ MES,
  data = smrt.mal@meta.data,
  exact = FALSE
)

str(model.wilcox)
```

```{r}
smrt.mal$Sample <- as.factor(smrt.mal$Sample)
levels(smrt.mal$Sample)

smrt.mal.meta <- smrt.mal@meta.data

# in smrt.mal.meta.list[[7]] it is all NA celltype_merge so we have to remove it
smrt.mal.meta <- smrt.mal.meta[!smrt.mal.meta$Sample == "BT85", ]
smrt.mal.meta$Sample <- droplevels(smrt.mal.meta$Sample)

smrt.mal.meta[smrt.mal.meta$Sample == "BT85", ]

levels(smrt.mal.meta$Sample)

smrt.mal.meta.list <- split(smrt.mal.meta, f = smrt.mal.meta$Sample)

wilcox.list <- list()
for (i in 1:length(smrt.mal.meta.list)) {
  print(i)
  wilcox.each <- wilcox.test(senescence1 ~ MES, data = smrt.mal.meta.list[[i]], exact = FALSE)
  wilcox.list <- append(wilcox.list, list(wilcox.each))
}



str(wilcox.list)

wilcox.list

for (i in 1:length(wilcox.list)) {
  print(wilcox.list[[i]]$p.value < 0.05)
}
```



```{r}
library(ggplot2)
# Basic box plot
p <- ggplot(smrt.mal@meta.data, aes(x = MES, y = senescence1, fill = MES)) +
  geom_boxplot(outlier.shape = NA)
p <- p + facet_wrap(~Sample)
p <- p + theme(
  axis.text.x.bottom = element_text(angle = 90, size = 5)
)
p
```

#Try annotation in several tools
```{r}
gbm <- readRDS(paste0("output/temp"))

library(scSorter)
# Scsorter
anno <- as.data.frame(read_xlsx("data/gbm_subtype_genelist_scsorter.xlsx", sheet = 2))

# Copied code >> and modified so that the each meta data is not the summation yet
rts.list <- list()
each.meta.list <- list()

gbm <- readRDS(paste0("output/temp"))
donor.id.list <- as.character(unique(gbm$Sample))
Idents(gbm) <- "Sample"

levels(Idents(gbm))

for (i in 1:length(donor.id.list)) {
  donor.id <- donor.id.list[i]
  each <- subset(gbm, idents = donor.id)
  # Just copy the code to here
  topgenes <- head(VariableFeatures(each), 2000)
  # expr = GetAssayData(each, assay = "SCT", slot = "data")
  expr <- GetAssayData(each, assay = "RNA", slot = "data")
  topgene_filter <- rowSums(as.matrix(expr)[topgenes, ] != 0) > ncol(expr) * .1
  topgenes <- topgenes[topgene_filter]
  picked_genes <- unique(c(anno$Marker, topgenes))
  expr <- expr[rownames(expr) %in% picked_genes, ]

  rts <- scSorter(expr, anno)

  # rts.list[[length(rts.list) + 1]] <- rts
  rts.list[[donor.id]] <- c(rts.list[[donor.id]], rts)

  each <- AddMetaData(each, rts$Pred_Type, col.name = "scSorter")
  each.meta <- each@meta.data[c("NAME", "Sample", "scSorter")]
  each.meta <- rownames_to_column(each.meta)
  each.meta.list[[donor.id]] <- c(each.meta.list[[donor.id]], each.meta)
}


str(each.meta.list)


# Demultiplex list
each.meta.df <- data.frame()
for (i in 1:length(each.meta.list)) {
  each.meta.df <- rbind(each.meta.df, (each.meta.list[[i]]))
}
colnames(each.meta.df)[1] <- "cell_id"

rownames(each.meta.df) <- each.meta.df$cell_id
each.meta.df
gbm <- AddMetaData(gbm, each.meta.df[4])
gbm@meta.data
DimPlot(gbm, group.by = "seurat_clusters")
```
```{r}
DimPlot(gbm, group.by = "scSorter")
```

```{r}
DimPlot(gbm, group.by = "celltype_merge")
DimPlot(gbm, group.by = "Sample")
```
```{r}
# try scSorter without splitting the cell
# Just copy the code to here
topgenes <- head(VariableFeatures(gbm), 2000)
# expr = GetAssayData(gbm, assay = "SCT", slot = "data")
expr <- GetAssayData(gbm, assay = "RNA", slot = "data")
topgene_filter <- rowSums(as.matrix(expr)[topgenes, ] != 0) > ncol(expr) * .1
topgenes <- topgenes[topgene_filter]
picked_genes <- unique(c(anno$Marker, topgenes))
expr <- expr[rownames(expr) %in% picked_genes, ]

rts <- scSorter(expr, anno)

gbm <- AddMetaData(gbm, rts$Pred_Type, col.name = "scSorter_all")
```
```{r}
p0 <- DimPlot(gbm, group.by = "scSorter_all", cols = c("AC" = "red", "MES" = "lightblue", "NPC" = "green", "OPC" = "orange", "Unknown" = "black")) + theme(legend.text = element_blank())
p1 <- DimPlot(gbm, group.by = "scSorter", cols = c("AC" = "red", "MES" = "lightblue", "NPC" = "green", "OPC" = "orange", "Unknown" = "black")) + labs(title = "scSorter_each")
p0 + p1
```

#markercount 
```{r}
# cannot done because no raw_count
```


```{r}
# sctype
lapply(c("dplyr", "Seurat", "HGNChelper"), library, character.only = T)


# load gene set preparation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/gene_sets_prepare.R")
# load cell type annotation function
source("https://raw.githubusercontent.com/IanevskiAleksandr/sc-type/master/R/sctype_score_.R")


# DB file
db_ <- "data/gbm_subtype_genelist_sctype_4equal.xlsx"
tissue <- "gbm"

# prepare gene sets
gs_list <- gene_sets_prepare(db_, tissue)
gs_list

es.max <- sctype_score(
  scRNAseqData = gbm[["RNA"]]@scale.data, scaled = TRUE,
  gs = gs_list$gs_positive, gs2 = NULL
)
cL_resutls <- do.call("rbind", lapply(unique(gbm@meta.data$seurat_clusters), function(cl) {
  es.max.cl <- sort(rowSums(es.max[, rownames(gbm@meta.data[gbm@meta.data$seurat_clusters == cl, ])]), decreasing = !0)
  head(data.frame(cluster = cl, type = names(es.max.cl), scores = es.max.cl, ncells = sum(gbm@meta.data$seurat_clusters == cl)), 10)
}))
sctype_scores <- cL_resutls %>%
  group_by(cluster) %>%
  top_n(n = 1, wt = scores)

sctype_scores$type[as.numeric(as.character(sctype_scores$scores)) < sctype_scores$ncells / 4] <- "Unknown"
gbm@meta.data$customclassif <- ""
for (j in unique(sctype_scores$cluster)) {
  cl_type <- sctype_scores[sctype_scores$cluster == j, ]
  gbm@meta.data$customclassif[gbm@meta.data$seurat_clusters == j] <- as.character(cl_type$type[1])
}
sctype <- as.data.frame(gbm$customclassif)
gbm$sctype_4equal <- sctype
```
```{r}
DimPlot(gbm, reduction = "umap", group.by = "sctype_4equal")
```
```{r}
Idents(gbm) <- "celltype_merge"
DimPlot(subset(gbm, idents = levels(gbm$celltype_merge)), reduction = "umap", group.by = "sctype_4equal", split.by = "celltype_merge", ncol = 2) + labs(subtitle = "sc-type; split by neftel_celltype_merge; group by sctype", caption = "unknown = /4 (default)") + theme(legend.title = element_text())
```


#10X of Neftel
```{r}
data.dir <- "data/neftel/"


neftel.10x.1 <- Read10X(data.dir = paste0(data.dir, "1/"))
neftel.10x.1 <- CreateSeuratObject(neftel.10x.1)
unique(neftel.10x.1$orig.ident)

neftel.10x.2 <- Read10X(data.dir = paste0(data.dir, "2/"))
neftel.10x.2 <- CreateSeuratObject(neftel.10x.2)
unique(neftel.10x.2$orig.ident)

neftel.10x <- merge(neftel.10x.1, y = neftel.10x.2)
```
```{r}
neftel.10x@meta.data
```

```{r}
# filter by MT, Count, feature
Idents(neftel.10x) <- "orig.ident"

neftel.10x[["percent.mt"]] <- PercentageFeatureSet(neftel.10x, pattern = "^MT-")
```

```{r}
VlnPlot(neftel.10x, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
Idents(neftel.10x) <- "orig.ident"
plot2 <- FeatureScatter(neftel.10x, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot3 <- FeatureScatter(neftel.10x, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2 + plot3
```

```{r}
library(scales)
ggplot(neftel.10x@meta.data, aes(x = nCount_RNA, y = percent.mt)) +
  geom_point(size = 0.2) +
  facet_wrap(~orig.ident, ncol = 4) +
  theme(
    axis.text.x = element_text(
      size = 5, angle = 90
    ),
    axis.text.y = element_text(
      size = 6
    )
  ) +
  scale_y_continuous(breaks = seq(0, 70, 10)) +
  scale_x_continuous(breaks = c(10000, 50000, 100000, 200000)) +
  geom_hline(yintercept = 15, linetype = "dashed", color = "red") +
  geom_vline(xintercept = 5e+04, linetype = "dashed", color = "red") +
  labs(title = "orig.ident")
```

```{r}
ggplot(neftel.10x@meta.data, aes(x = nCount_RNA, y = nFeature_RNA)) +
  geom_point(size = 0.2) +
  facet_wrap(~donor_id, ncol = 4) +
  theme(
    axis.text.x = element_text(
      size = 5, angle = 90
    ),
    axis.text.y = element_text(
      size = 5
    )
  ) +
  scale_y_continuous(breaks = seq(0, 12500, 1000)) +
  # scale_x_continuous(breaks=c(10000,50000,100000,200000))
  geom_hline(yintercept = 200, linetype = "dashed", color = "red") +
  geom_hline(yintercept = 12000, linetype = "dashed", color = "red") +
  geom_vline(xintercept = 10000, linetype = "dashed", color = "red") +
  labs(title = donor.group)
```

```{r}
# Filter the same way as neftel
neftel.10x.list <- SplitObject(neftel.10x, split.by = "orig.ident")
neftel.10x.list <- lapply(X = neftel.10x.list, FUN = function(x) {
  mean.feature <- mean(x$nFeature_RNA)
  x <- subset(x, subset = nFeature_RNA > (0.5 * mean.feature) & nFeature_RNA < (2 * mean.feature) & percent.mt < 15)
})
```
```{r}
x <- 0
for (i in neftel.10x.list) {
  x <- x + nrow(i@meta.data)
}
x
nrow(neftel.10x@meta.data)
1 - x / nrow(neftel.10x@meta.data)
```
```{r}
library(scCustomize)
neftel.10x <- Merge_Seurat_List(
  neftel.10x.list,
  add.cell.ids = NULL,
  merge.data = TRUE
)
```
```{r}
VlnPlot(neftel.10x, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
```{r}
non.malig <- list(
  mac = c("CD14", "AIF1", "FCER1G", "FCGR3A", "TYROBP", "CSF1R"),
  tcell = c("CD2", "CD3D", "CD3E", "CD3G"),
  oli = c("MBP", "TF", "PLP1", "MAG", "MOG", "CLDN11")
)
names(non.malig)
```

```{r}
neftel.10x <- AddModuleScore(neftel.10x, features = non.malig, name = names(non.malig), nbin = 30, ctrl = 100)
```

```{r}
neftel.10x@meta.data
neftel.10x.mal <- !neftel.10x$mac1 >= 0.5
table(neftel.10x)
hist(neftel.10x$tcell2)
hist(neftel.10x$oli3)
```


```{r}
neftel.10x.mal <- subset(x = neftel.10x, subset = oli3 < 3 & mac1 < 0.5)
```

```{r}
neftel.10x.mal <- NormalizeData(neftel.10x.mal, normalization.method = "LogNormalize", scale.factor = 10000)
all.genes <- rownames(neftel.10x.mal)
neftel.10x.mal <- ScaleData(neftel.10x.mal, features = all.genes)
neftel.10x.mal <- FindVariableFeatures(neftel.10x.mal, selection.method = "vst", nfeatures = 2000)

neftel.10x.mal <- RunPCA(neftel.10x.mal, features = VariableFeatures(object = neftel.10x.mal))
neftel.10x.mal <- FindNeighbors(neftel.10x.mal, dims = 1:10)
neftel.10x.mal <- FindClusters(neftel.10x.mal, resolution = 0.5)

neftel.10x.mal <- RunUMAP(neftel.10x.mal, dims = 1:10)

DimPlot(neftel.10x.mal, reduction = "umap", group.by = "orig.ident")
```

```{r}
# Add score for subtype
library(UCell)

# This list is 4 without reducing
signatures <- read_xlsx("data/gbm_subtype_genelist_original.xlsx", sheet = 4)

colnames(signatures)[5] <- "G1_S"
colnames(signatures)[6] <- "G2_M"

signatures <- as.list(signatures)
signatures <- lapply(signatures, na.omit)
signatures <- signatures[-c(5, 6)]

neftel.10x.mal <- AddModuleScore_UCell(neftel.10x.mal, features = signatures)
neftel.10x.mal <- AddModuleScore(neftel.10x.mal, features = signatures, name = names(signatures))
```

```{r}
FeaturePlot(neftel.10x.mal, features = c("MES_UCell", "NPC_UCell", "OPC_UCell", "AC_UCell"), label = FALSE)
```
```{r}
FeaturePlot(neftel.10x.mal, features = c("MES1", "NPC4", "OPC3", "AC2"))
```
```{r}
library(ggpubr)
ggviolin(neftel.10x.mal@meta.data, y = c("MES_UCell", "NPC_UCell", "OPC_UCell", "AC_UCell"), merge = TRUE) + facet_wrap("orig.ident", ncol = 5) + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + labs(title = "cellular state signatures", subtitle = "control & 2 runs; signatures = 4 signatures (merge without filter)", caption = "AddModuleScore(seurat)")
```

```{r}
library(ggpubr)
ggviolin(neftel.10x.mal@meta.data, y = c("MES1", "NPC4", "OPC3", "AC2"), merge = TRUE) + facet_wrap("orig.ident", ncol = 5) + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + labs(title = "cellular state signatures", subtitle = "control & 2 runs; signatures = 4 signatures (merge without filter)", caption = "AddModuleScore(seurat)")
```












